Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-02 reply_to_reviewer2.tex

\paragraph{Response}

\hrulefill

\paragraph{3. Reviewer} \textit{Line 184: L-BFGS-B. It might be good to spell it out for clarity.}

\paragraph{Response} Done. We updated the revised manuscript as follows (\red{line XX, p. YY}):

`` 
$\dots$ solved using the \revision{L-BFGS-B (limited-memory Broyden--Fletcher--Goldfarb--Shanno with bounds support) algorithm [40]}, which is an $\dots$
''

@@

\paragraph{Response} Done.

\hrulefill

\paragraph{5. Reviewer} \textit{Line 211: corresponding notation for dimensions should be provided, or reference to Table 1 should be given right after 4.8x2.4x1}

\paragraph{Response} Done. In addition, we changed the notation of the domain height from $H$ to $L_z$ for consistency. The manuscript is revised as follows (\red{line XX, p. YY}):

`` 
$\dots$ has a \revision{total dimension of $L_x \times L_y \times L_z = 4.8 \times 2.4 \times 1 $ km$^3$, with $L_x, L_y$, and $L_z$ the streamwise, transversal, and vertical extent of the domain respectively. This domain is discretized on a grid of $N_x \times N_y \times N_z = 256 \times 128 \times 128$ gridpoints.} The final $\dots$
''

### 2018-01-02 MuntersMeyers_DynamicStrategiesControl_revision_WM01.tex
R2C4
R2C5

\begin{equation}
\nabla \Jtilde \equiv 
\begin{bmatrix*}[l]
\partial \Jtilde / \partial \bs{C}_T' \\
\partial \Jtilde / \partial \bs{\omega} 
\end{bmatrix*} = 
\begin{bmatrix}
- \bs{\sigma}\\
- \bs{\eta}
\end{bmatrix} \label{eq:problem_gradient}.
\end{equation}
The adjoint gradient expression shown above is derived and compared to a finite difference gradient approximation in Appendix A.

@@

\subsection{Simulation setup}
\noindent The simulation domain has a total dimension of $4.8 \times 2.4 \times 1 $ km$^3$, and is discretized on a grid of $256 \times 128 \times 128$ gridpoints. The final $10\%$ of the streamwise extent of the domain is used as a fringe region to impose inflow conditions. All cases apply periodic boundary conditions on transversal domain boundaries.

Two different sets of simulations are considered. The first set is subjected to a steady and uniform inflow with $U_\infty = 8$ m s$^{-1}$. In this case, the top and bottom boundaries are treated with a symmetry condition, and the turbines are placed at the vertical mid-plane $z_h/H = 0.5$ to minimize any influence of vertical domain boundaries. The second set applies unsteady turbulent inflow conditions, which are generated in a pressure-driven precursor simulation of an unperturbed turbulent boundary layer with shifted periodic boundary conditions \cite{munters2016shifted}, driven by a pressure gradient of $\partial_x p_\infty/\rho = 2.5 \times 10^{-4}$ m s$^{-2}$, resulting in a freestream wind speed at hub height slightly higher than 8 m s$^{-1}$. In this second case, the top boundary is treated with a stress-free condition, and a high-Reynolds number wall model is used at the bottom. The turbines are placed at a height of $z_h/H = 0.1$ from the wall. The wall model adds a wall stress consistent with a rough-wall logarithmic boundary layer profile with roughness length $z_0/H = 10^{-4}$, resulting in a inflow turbulence intensity of about 8\% at hub height.

@@

\begin{table}
	\caption{Setup parameters for optimal dynamic induction and yaw control cases \label{tab:case_definition}}
	\centering
	\begin{tabular}{llccc}
		\toprule
		\multicolumn{5}{l}{\textbf{Simulation parameters}}\\
		\midrule
		Domain size  			& \multicolumn{4}{l}{$L_x \times L_y \times H = 4.8 \times 2.4 \times 1$ km$^3$}  \\ 
		Turbine dimensions  		& \multicolumn{4}{l}{$D = 0.1H = 100$ m}\\ 
		Turbine spacing  		& \multicolumn{4}{l}{$S_x = 6D, \quad S_y = 6D$}\\
		Windfarm layout 		& \multicolumn{4}{l}{$N_t = 16 $ turbines = 4 rows $\times$ 4 columns} \\ 
		Grid size 			& \multicolumn{4}{l}{$N_x \times N_y \times N_z = 256 \times 128 \times 128$}\\
		Cell size 			& \multicolumn{4}{l}{$\Delta_x \times \Delta_y \times \Delta_z = 18.75 \times 18.75 \times 7.8$ m$^3$}\\
		Time step 			& \multicolumn{4}{l}{$\Delta t = 0.75$ s}\\		
		& & & & \\	
		\multicolumn{5}{l}{\textit{Uniform inflow}}\\
		Hub height & \multicolumn{4}{l}{$z_h/H$ = 0.5}\\
		Inflow velocity & \multicolumn{4}{l}{$U_\infty$ = 8 m s$^{-1}$}\\
		& & & & \\			
		\multicolumn{5}{l}{\textit{Turbulent inflow}}\\
		Hub height & \multicolumn{4}{l}{$z_h/H$ = 0.1}\\		
		Precursor pressure gradient  	& \multicolumn{4}{l}{$ \partial_x p_\infty = 2.5 \times 10^{-4}$ m s$^{-2}$}  \\ 
		Surface roughness  &  \multicolumn{4}{l}{$z_0 = 10^{-4}H = 0.1$ m}\\ 		
		& & & & \\	
		\toprule
		\multicolumn{5}{l}{\textbf{Optimization parameters}}\\
		\midrule
		Optimization method		& \multicolumn{4}{l}{L--BFGS--B} \\
		Hessian correction pairs	& \multicolumn{4}{l}{$m = 5$} \\
		BFGS iterations 		& \multicolumn{4}{l}{$N_{it} = 80$ ($\approx 160$ PDE)} \\
		Optimization time window	& \multicolumn{4}{l}{$T = 300$ s}\\
		Flow advancement time window 	& \multicolumn{4}{l}{$T_A = 150$ s}\\
		Total operation time         	& \multicolumn{4}{l}{$T_{tot} = 900$ s (6 windows)}\\
		& & & & \\	
		\toprule
		\textbf{Cases} & \  & $C_{T,\text{min}}'$ & $C_{T,\text{max}}'$ & $\omega_{\text{max}}$ [$^\circ/s$] \\ 
		\toprule
		Reference & (R)    &  2 & 2 & 0   \\ 
		Yaw control & (Y)   &  2 & 2 & \textbf{0.3} \\ 
		Fast yaw control & (Y5)   &  2 & 2 &\textbf{5}   \\ 
		Underinduction control & (I2)    &  \textbf{0} & 2 & 0\\
		Overinduction control & (I3)    &  \textbf{0} & \textbf{3} & 0\\ 
		Underinduction + yaw control & (I2Y)   &  \textbf{0} & 2 & \textbf{0.3} \\ 
		Overinduction + yaw control & (I3Y)   &  \textbf{0} & \textbf{3} & \textbf{0.3} \\ 
		\bottomrule
	\end{tabular} 
\end{table}

>>>>

\begin{equation}
\nabla \Jtilde \equiv 
\begin{bmatrix*}[l]
\partial \Jtilde / \partial \bs{C}_T' \\
\partial \Jtilde / \partial \bs{\omega} 
\end{bmatrix*} = 
\begin{bmatrix}
- \bs{\sigma}\\
- \bs{\eta}
\end{bmatrix} \label{eq:problem_gradient}.
\end{equation}

@@

\subsection{Simulation setup}
\noindent The simulation domain has a \revision{total dimension of $L_x \times L_y \times L_z = 4.8 \times 2.4 \times 1 $ km$^3$, with $L_x, L_y$, and $L_z$ the streamwise, transversal, and vertical extent of the domain respectively. This domain is discretized on a grid of $N_x \times N_y \times N_z = 256 \times 128 \times 128$ gridpoints.} The final $10\%$ of the streamwise extent of the domain is used as a fringe region to impose inflow conditions. All cases apply periodic boundary conditions on transversal domain boundaries.

Two different sets of simulations are considered. The first set is subjected to a steady and uniform inflow with $U_\infty = 8$ m s$^{-1}$. In this case, the top and bottom boundaries are treated with a symmetry condition, and the turbines are placed at the vertical mid-plane $z_h/L_z = 0.5$ to minimize any influence of vertical domain boundaries. The second set applies unsteady turbulent inflow conditions, which are generated in a pressure-driven precursor simulation of an unperturbed turbulent boundary layer with shifted periodic boundary conditions \cite{munters2016shifted}, driven by a pressure gradient of $\partial_x p_\infty/\rho = 2.5 \times 10^{-4}$ m s$^{-2}$, resulting in a freestream wind speed at hub height slightly higher than 8 m s$^{-1}$. In this second case, the top boundary is treated with a stress-free condition, and a high-Reynolds number wall model is used at the bottom. The turbines are placed at a height of $z_h/L_z = 0.1$ from the wall. The wall model adds a wall stress consistent with a rough-wall logarithmic boundary layer profile with roughness length $z_0/L_z = 10^{-4}$, resulting in a inflow turbulence intensity of about 8\% at hub height.

@@

\begin{table}
	\caption{Setup parameters for optimal dynamic induction and yaw control cases \label{tab:case_definition}}
	\centering
	\begin{tabular}{llccc}
		\toprule
		\multicolumn{5}{l}{\textbf{Simulation parameters}}\\
		\midrule
		Domain size  			& \multicolumn{4}{l}{$L_x \times L_y \times L_z = 4.8 \times 2.4 \times 1$ km$^3$}  \\ 
		Turbine dimensions  		& \multicolumn{4}{l}{$D = 0.1L_z = 100$ m}\\ 
		Turbine spacing  		& \multicolumn{4}{l}{$S_x = 6D, \quad S_y = 6D$}\\
		Windfarm layout 		& \multicolumn{4}{l}{$N_t = 16 $ turbines = 4 rows $\times$ 4 columns} \\ 
		Grid size 			& \multicolumn{4}{l}{$N_x \times N_y \times N_z = 256 \times 128 \times 128$}\\
		Cell size 			& \multicolumn{4}{l}{$\Delta_x \times \Delta_y \times \Delta_z = 18.75 \times 18.75 \times 7.8$ m$^3$}\\
		Time step 			& \multicolumn{4}{l}{$\Delta t = 0.75$ s}\\		
		& & & & \\	
		\multicolumn{5}{l}{\textit{Uniform inflow}}\\
		Hub height & \multicolumn{4}{l}{$z_h/L_z$ = 0.5}\\
		Inflow velocity & \multicolumn{4}{l}{$U_\infty$ = 8 m s$^{-1}$}\\
		& & & & \\			
		\multicolumn{5}{l}{\textit{Turbulent inflow}}\\
		Hub height & \multicolumn{4}{l}{$z_h/L_z$ = 0.1}\\		
		Precursor pressure gradient  	& \multicolumn{4}{l}{$ \partial_x p_\infty = 2.5 \times 10^{-4}$ m s$^{-2}$}  \\ 
		Surface roughness  &  \multicolumn{4}{l}{$z_0 = 10^{-4}H = 0.1$ m}\\ 		
		& & & & \\	
		\toprule
		\multicolumn{5}{l}{\textbf{Optimization parameters}}\\
		\midrule
		Optimization method		& \multicolumn{4}{l}{L--BFGS--B} \\
		Hessian correction pairs	& \multicolumn{4}{l}{$m = 5$} \\
		BFGS iterations 		& \multicolumn{4}{l}{$N_{it} = 80$ ($\approx 160$ PDE)} \\
		Optimization time window	& \multicolumn{4}{l}{$T = 300$ s}\\
		Flow advancement time window 	& \multicolumn{4}{l}{$T_A = 150$ s}\\
		Total operation time         	& \multicolumn{4}{l}{$T_{tot} = 900$ s (6 windows)}\\
		& & & & \\	
		\toprule
		\textbf{Cases} & \  & $C_{T,\text{min}}'$ & $C_{T,\text{max}}'$ & $\omega_{\text{max}}$ [$^\circ/s$] \\ 
		\toprule
		Reference & (R)    &  2 & 2 & 0   \\ 
		Yaw control & (Y)   &  2 & 2 & \textbf{0.3} \\ 
		Fast yaw control & (Y5)   &  2 & 2 &\textbf{5}   \\ 
		Underinduction control & (I2)    &  \textbf{0} & 2 & 0\\
		Overinduction control & (I3)    &  \textbf{0} & \textbf{3} & 0\\ 
		Underinduction + yaw control & (I2Y)   &  \textbf{0} & 2 & \textbf{0.3} \\ 
		Overinduction + yaw control & (I3Y)   &  \textbf{0} & \textbf{3} & \textbf{0.3} \\ 
		\bottomrule
	\end{tabular} 
\end{table}
